Near Optimal Behavior with Approximate State Abstraction

### 2016-02-04 paper/theory_results.tex
Added some notes about Qmax
Added some notes

% Theorem 1
\begin{thm}
There exist at least four classes of approximate state-aggregation functions, $\ep_{Q^*}$, $\ep_{\text{model}}$, $\ep_{\text{bolt}}$ and $\ep_{\text{mult}}$ for which the optimal policy in the abstract \ac{MDP}, applied to the ground \ac{MDP}, has suboptimality bounded polynomially in $\epsilon$:
\begin{equation}
\forall_{s \in \mathcal{S}_G}: | V_G^{\pi_{GA}}(s) - V_G^{\pi_G^*}(s) | \leq poly(\epsilon).
\end{equation}
\end{thm}
\enote{Axe the absolute values here and elsewhere}
\enote{Unify the use of $\forall$ and ``:"s}

@@

% Q* Lemma
We consider an approximate version of the $Q^*$-irrelevance abstraction. An approximate $Q$ function has the same form as Equation~\ref{eq:phi_f}:
\begin{equation}
\ep_{Q^*}(s_1) = \ep_{Q^*}(s_2) \rightarrow \forall_a \left|Q_G^*(s_1, a) - Q_G^*(s_2, a)\right| \leq \epsilon.
\end{equation}

% Q^* Lemma
\begin{lma}
\label{lma:Q*}
When $\ep_{Q^*}$ is used to create the abstract state space $\mcS_A$:
\begin{equation}
\forall_{s \in \mcS_G}: | V^{\pi_G^*}(s) - V^{\pi_{GA}}(s) | \leq \frac{2\epsilon}{(1-\gamma)^2}.
\end{equation}
\end{lma}

@@

\begin{multline*}
Q_T(s_A, a) = \sum_{g \in G(s_A)}w(g)* \\ 
 \left[ R_0(g,a) + \gamma \sum_{g' \in \mathcal{S}_G} T_0(g,a,g') \max_{a'} Q_{T-1}(g', a')      \right]
\end{multline*}
Applying our inductive hypothesis yields,
\begin{multline*}
\leq \sum_{g \in G(s_A)}w(g) * \\ \left[ R_G(g,a) + \gamma \sum_{g' \in \mathcal{S}_G} T_G(g,a,g') \max_{a'}(Q_G(g', a') + \sum_{t=0}^{T-2} \epsilon \gamma^t)      \right]
\end{multline*}
%&\leq \gamma\sum_{t=0}^{T-2} \epsilon \gamma^t + \sum_{g \in X(s_A)}w(g)\left[ R_0(g,a) + \gamma \sum_{g' \in \mathcal{S}_G} T_0(g,a,g') \max_{a'}Q_0      \right]\\
Then:
\begin{equation*}
\leq \gamma\sum_{t=0}^{T-2} \epsilon \gamma^t + \sum_{g \in X(s_A)}\left[ w(g)\ Q_G(g,a)\right]
\end{equation*}
Since all aggregated states have Q-values within $\epsilon$ of one another:
\begin{align*}
&\leq \gamma\sum_{t=0}^{T-2} \epsilon \gamma^t + \epsilon + Q_G(s_G, a)\\
&\leq \gamma\sum_{t=0}^{T-1} \epsilon \gamma^t + Q_G(s_G, a)
\end{align*}
Thus,
\begin{equation*}
\left| Q_{T}(s_A, a) - Q_G(s_G,a) \right| \leq \gamma\sum_{t=0}^{T-1}\epsilon \gamma^t
\end{equation*}

@@

% Subsection: Model Similarity
\subsection{Model Similarity: $\ep_{model}$}
\label{sec:model}

Inspired by the model-irrelevance abstraction, we consider the approximate case. We let $\epM$ define a type of abstraction that, for fixed $\epsilon$, satisfies:
\begin{multline}
\epM(s_1) = \epM(s_2) \rightarrow \\
\forall_a \left| \mcR_G(s_1, a) - \mcR_G(s_2, a)\right| \leq \epsilon \wedge\ \\
\forall_{s_A \in \mcS_A} \left|\sum_{{s_G}' \in G(s_A)} \left[\mcT_G(s_1, a, {s_G}') - \mcT_G(s_2, a,{s_G}')\right] \right| \leq \epsilon
\end{multline}

%Model Lemma
\begin{lma}
\label{lma:model}
When $\mcS_A$ is created using a function of the $\ep_{model}$ type:
\begin{equation}
\forall_{s \in \mcS_G} : | V_G(s) - V^{\pi_{GA}}(s) | \leq \frac{2\epsilon + 2\gamma((|\mcS_G|-1)\epsilon)}{(1-\gamma)^3}
\end{equation}
\end{lma}

@@

@@

We let $\phi_{bolt}$ define a family of abstractions that, for fixed $\epsilon$ satisfies:
\begin{multline}
\ep_{bolt}(s_1) = \ep_{bolt}(s_2) \rightarrow \\
\forall_{a} \left|\frac{e^{Q^*(s_1,a)}}{\sum_b e^{Q^*(s_1,b)}} - \frac{e^{Q^*(s_2,a)}}{\sum_b e^{Q^*(s_2,b)}}\right| \leq \epsilon
\label{eq:phi_bolt}
\end{multline}

@@

\begin{lma} When $S_A$ is created using a function of the $\ep_{bolt}$ type, for some non-negative constant $k \in \mathbb{R}$::
\begin{equation}
\forall_{s \in S_M} : | V^{\pi^*_G}(s) - V^{\pi^*_{A}}(s) | \leq \frac{2\epsilon\left(\frac{|\mathcal{A}|}{1-\gamma} + k\epsilon + k\right)}{(1-\gamma)^2}
\end{equation}
\label{lma:bolt_lemma}
\end{lma}

@@

We consider a multinomial distribution over $Q^*$ for similar reasons as the Boltzmann. We let $\phi_{mult}$ define a type of abstraction that, for fixed $\epsilon$ satisfies.
\begin{multline}
\ep_{mult}(s_1) = \ep_{mult}(s_2) \rightarrow \\
\forall_{a} \left|\frac{Q^*(s_1,a)}{\sum_b Q^*(s_1,b)} - \frac{Q^*(s_1,a)}{\sum_b Q^*(s_1,b)}\right| \leq \epsilon
\end{multline}

>>>>

% Theorem 1
\begin{thm}
There exist at least four classes of approximate state-aggregation functions, $\ep_{Q^*}$, $\ep_{\text{model}}$, $\ep_{\text{bolt}}$ and $\ep_{\text{mult}}$ for which the optimal policy in the abstract \ac{MDP}, applied to the ground \ac{MDP}, has suboptimality bounded polynomially in $\epsilon$:
\begin{equation}
\forall_{s \in \mathcal{S}_G}: | V_G^{\pi_{GA}}(s) - V_G^{\pi_G^*}(s) | \leq poly(\epsilon).
\end{equation}
\end{thm}
\enote{Axe the absolute values here and elsewhere}

\enote{Unify the use of $\forall$ and ``:"s}

\enote{Add for 0-1 normalized reward}

\enote{Add note relating this all to Lihong for eps=0 and cut down the Lihong section}

@@

% Q* Lemma -- lemma 1
We consider an approximate version of the $Q^*$-irrelevance abstraction.

\bdefn{$\epQ$}
An approximate $Q$ function has the same form as Equation~\ref{eq:phi_f}:
\begin{equation}
\ep_{Q^*}(s_1) = \ep_{Q^*}(s_2) \rightarrow \forall_a \left|Q_G^*(s_1, a) - Q_G^*(s_2, a)\right| \leq \epsilon.
\end{equation}
\edefn

% Q^* Lemma
\begin{lma}
\label{lma:Q*}
When $\ep_{Q^*}$ is used to create the abstract state space $\mcS_A$:
\begin{equation}
\forall_{s \in \mcS_G}: | V_G^{\pi_G^*}(s) - V_G^{\pi_{GA}}(s) | \leq \frac{2\epsilon}{(1-\gamma)^2}.
\end{equation}
\end{lma}

@@

\begin{multline*}
Q_T(s_A, a) = \sum_{g \in G(s_A)}w(g)* \\ 
 \left[ R_0(g,a) + \gamma \sum_{g' \in \mathcal{S}_G} T_0(g,a,g') \max_{a'} Q_{T-1}(g', a')      \right]
\end{multline*}
Applying our inductive hypothesis yields,
\begin{multline*}
\leq \sum_{g \in G(s_A)}w(g) * \\ \left[ R_G(g,a) + \gamma \sum_{g' \in \mathcal{S}_G} T_G(g,a,g') \max_{a'}(Q_G(g', a') + \sum_{t=0}^{T-2} \epsilon \gamma^t)      \right]
\end{multline*}
%&\leq \gamma\sum_{t=0}^{T-2} \epsilon \gamma^t + \sum_{g \in X(s_A)}w(g)\left[ R_0(g,a) + \gamma \sum_{g' \in \mathcal{S}_G} T_0(g,a,g') \max_{a'}Q_0      \right]\\
Then:
\begin{equation*}
\leq \gamma\sum_{t=0}^{T-2} \epsilon \gamma^t + \sum_{g \in G(s_A)}\left[ w(g)\ Q_G(g,a)\right]
\end{equation*}
Since all aggregated states have Q-values within $\epsilon$ of one another:
\begin{align*}
&\leq \gamma\sum_{t=0}^{T-2} \epsilon \gamma^t + \epsilon + Q_G(s_G, a)\\
&\leq \gamma\sum_{t=0}^{T-1} \epsilon \gamma^t + Q_G(s_G, a)
\end{align*}
Thus,
\begin{equation*}
\left| Q_{T}(s_A, a) - Q_G(s_G,a) \right| \leq \gamma\sum_{t=0}^{T-1}\epsilon \gamma^t
\end{equation*}

@@

% Subsection: Model Similarity -- lemma 2
\subsection{Model Similarity: $\ep_{model}$}
\label{sec:model}

Inspired by the model-irrelevance abstraction, we consider the approximate case. 
\bdefn{$\epM$}
We let $\epM$ define a type of abstraction that, for fixed $\epsilon$, satisfies:
\begin{multline}
\epM(s_1) = \epM(s_2) \rightarrow \\
\forall_a \left| \mcR_G(s_1, a) - \mcR_G(s_2, a)\right| \leq \epsilon \wedge\ \\
\forall_{s_A \in \mcS_A} \left|\sum_{{s_G}' \in G(s_A)} \left[\mcT_G(s_1, a, {s_G}') - \mcT_G(s_2, a,{s_G}')\right] \right| \leq \epsilon
\end{multline}
\edefn

%Model Lemma
\begin{lma}
\label{lma:model}
When $\mcS_A$ is created using a function of the $\ep_{model}$ type:
\begin{equation}
\forall_{s \in \mcS_G} : | V_G(s) - V_G^{\pi_{GA}}(s) | \leq \frac{2\epsilon + 2\gamma((|\mcS_G|-1)\epsilon)}{(1-\gamma)^3}
\end{equation}
\end{lma}

@@

\enote{Do we need to state what Q max is and that it is upper bounded by 1 over 1 minus gamma or is this common knowledge?}

@@

\bdefn{$\phi_{bolt}$}
We let $\phi_{bolt}$ define a family of abstractions that, for fixed $\epsilon$ satisfies:
\begin{multline}
\ep_{bolt}(s_1) = \ep_{bolt}(s_2) \rightarrow \\
\forall_{a} \left|\frac{e^{Q_G^*(s_1,a)}}{\sum_b e^{Q_G^*(s_1,b)}} - \frac{e^{Q_G^*(s_2,a)}}{\sum_b e^{Q_G^*(s_2,b)}}\right| \leq \epsilon
\label{eq:phi_bolt}
\end{multline}
\edefn

@@

\begin{lma} When $S_A$ is created using a function of the $\ep_{bolt}$ type, for some non-negative constant $k \in \mathbb{R}$::
\begin{equation}
\forall_{s \in \mcS_G} : | V^{\pi^*_G}(s) - V^{\pi_{GA}}(s) | \leq \frac{2\epsilon\left(\frac{|\mathcal{A}|}{1-\gamma} + k\epsilon + k\right)}{(1-\gamma)^2}
\end{equation}
\label{lma:bolt_lemma}
\end{lma}

@@

We consider a multinomial distribution over $Q^*$ for similar reasons as the Boltzmann.

\bdefn{$\ep_{mult}$}
We let $\phi_{mult}$ define a type of abstraction that, for fixed $\epsilon$ satisfies.
\begin{multline}
\ep_{mult}(s_1) = \ep_{mult}(s_2) \rightarrow \\
\forall_{a} \left|\frac{Q^*(s_1,a)}{\sum_b Q^*(s_1,b)} - \frac{Q^*(s_1,a)}{\sum_b Q^*(s_1,b)}\right| \leq \epsilon
\end{multline}
\edefn
